### 2015-04-15 Manuscript/Tables/table_literature.tex
Minor edits

\begin{longtable}{p{6cm}llllll}
\hline

\textbf{Paper}	&	Year	&	Journal	&	Taxon	&	Ntips	&	Use in paper	&	Stats/R package	\\
\hline
(Hansen and Orzack 2005)	&	2005	&	Evolution	&	insects	&	15	&	other	&	OUCH precursor?	\\
(Edwards and Donoghue 2006) 	&	2006	&	Am Nat	&	plants	&	12	&	ancestral state reconstruction	&	COMPARE	\\
(Gvoždík and Damme 2006)	&	2006	&	Evolution	&	amphibians	&	10	&	phylogenetic correction	&	COMPARE	\\
(Halsey et al. 2006)	&	2006	&	Am Nat	&	birds/mammals	&	90	&	phylogenetic correction	&	Custom code	\\
(Ives and Godfray 2006)	&	2006	&	Am Nat	&	insects	&	8	&	phylogenetic signal	&	MATLAB	\\
(Valiente-Banuet et al. 2006)	&	2006	&	PNAS	&	plants	&	47	&	model of evolution	&	? (OUCH)	\\
(Clabaut et al. 2007)	&	2007	&	Evolution	&	fish	&	45	&	phylogenetic correction	&	APE	\\
(Gomez and Théry 2007)	&	2007	&	Am Nat	&	birds	&	40	&	model of evolution	&	OUCH	\\
(Hipp 2007)	&	2007	&	Evolution	&	plants	&	53	&	model of evolution	&	BayesTraits	\\
(Rezende et al. 2007)	&	2007	&	Nature	&	plants/insects	&	?	&	phylogenetic signal	&	? (OUCH)	\\
(Spoor et al. 2007)	&	2007	&	PNAS	&	mammals	&	210	&	phylogenetic correction	&	PDAP	\\
(Stuart‐Fox et al. 2007)	&	2007	&	Am Nat	&	reptiles	&	21	&	phylogenetic correction	&	COMPARE	\\
(Buchwalter et al. 2008)	&	2008	&	PNAS	&	insects	&	21	&	phylogenetic correction	&	MATLAB	\\
(Dumont and Payseur 2008)	&	2008	&	Evolution	&	mammals	&	13	&	model of evolution	&	OUCH	\\
(Hansen et al. 2008)	&	2008	&	Evolution	&	mammals	&	105	&	other	&	SLOUCH	\\
(Smith et al. 2008)	&	2008	&	Evolution	&	plants	&	15	&	phylogenetic correction	&	APE	\\
(Warne and Charnov 2008)	&	2008	&	Am Nat	&	reptiles	&	71	&	phylogenetic correction	&	MATLAB	\\
(Adams et al. 2009)	&	2009	&	PRSB	&	amphibians	&	10	&	model of evolution	&	OUCH	\\
(Addison et al. 2009)	&	2009	&	PRSB	&	birds	&	23	&	phylogenetic correction	&	MATLAB	\\
(Agrawal et al. 2009)	&	2009	&	PNAS	&	plants	&	53	&	model of evolution	&	GEIGER	\\
(Bergmann et al. 2009)	&	2009	&	Evolution	&	reptiles	&	38	&	model of evolution	&	OUCH	\\
(Collar et al. 2009)	&	2009	&	Evolution	&	fish	&	29	&	model of evolution	&	OUCH	\\
(Gonzalez-Voyer et al. 2009)	&	2009	&	PRSB	&	fish	&	39	&	phylogenetic correction	&	APE/COMPARE	\\
(Goodman et al. 2009)	&	2009	&	Evolution	&	reptiles	&	20	&	phylogenetic correction	&	COMPARE	\\
(Huey et al. 2009)	&	2009	&	PRSB	&	reptiles	&	70	&	phylogenetic correction	&	MATLAB	\\
(Kozak et al. 2009)	&	2009	&	Evolution	&	amphibians	&	184	&	model of evolution	&	GEIGER	\\
(Labra et al. 2009)	&	2009	&	Am Nat	&	reptiles	&	83	&	other	&	SLOUCH	\\
(Rezende et al. 2009)	&	2009	&	ELE	&	fish	&	116	&	model of evolution	&	PDTREE	\\
(Swanson and Garland 2009)	&	2009	&	Evolution	&	birds	&	44	&	phylogenetic correction	&	MATLAB	\\
(Van Buskirk 2009)	&	2009	&	JEB	&	amphibians	&	82	&	other	&	SLOUCH	\\
(Burbrink and Pyron 2010)	&	2010	&	Evolution	&	reptiles	&	29	&	model of evolution	&	GEIGER	\\
(Cooper and Purvis 2010)	&	2010	&	Am Nat	&	mammals	&	45	&	model of evolution	&	GEIGER	\\
(Edwards and Smith 2010)	&	2010	&	PNAS	&	plants	&	300	&	model of evolution	&	OUCH	\\
(Harmon et al. 2010)	&	2010	&	Evolution	&	multiple	&	17	&	model of evolution	&	GEIGER	\\
(Helmus et al. 2010)	&	2010	&	ELE	&	zooplankton	&	15	&	phylogenetic correction	&	?	\\
(Kalinka et al. 2010)	&	2010	&	Nature	&	insects	&	6	&	model of evolution	&	OUCH	\\
(Kozak and Wiens 2010b)	&	2010	&	Am Nat	&	amphibians	&	84	&	model of evolution/	&	GEIGER/	\\
	&		&		&		&		&	ancestral state reconstruction	&	OUCH	\\
(Kozak and Wiens 2010a)	&	2010	&	ELE	&	amphibians	&	11	&	model of evolution	&	OUCH	\\
(Ord et al. 2010)	&	2010	&	Evolution	&	reptiles	&	16	&	other	&	SLOUCH	\\
(Price et al. 2010)	&	2010	&	Evolution	&	fish	&	122	&	model of evolution	&	GEIGER	\\
(Slater et al. 2010)	&	2010	&	PRSB	&	mammals	&	84	&	model of evolution	&	OUCH	\\
(Angielczyk et al. 2011)	&	2011	&	Evolution	&	reptiles	&	8	&	model of evolution	&	OUCH	\\
(Benesh et al. 2011)	&	2011	&	Evolution	&	helminths	&	310	&	model of evolution	&	OUCH	\\
(Collar et al. 2011)	&	2011	&	Evolution	&	reptiles	&	37	&	model of evolution	&	Brownie	\\
(Derryberry et al. 2011)	&	2011	&	Evolution	&	birds	&	285	&	model of evolution	&	GEIGER	\\
(Galvan and Moller 2011)	&	2011	&	JEB	&	birds	&	323	&	phylogenetic correction	&	COMPARE	\\
(Gonzalez-Voyer and Kolm 2011)	&	2011	&	JEB	&	fish	&	49	&	model of evolution	&	GEIGER	\\
(Ord et al. 2011)	&	2011	&	Evolution	&	multiple	&	23	&	phylogenetic signal	&	SLOUCH	\\
(Oufiero et al. 2011) 	&	2011	&	Evolution	&	reptiles	&	106	&	phylogenetic correction	&	SLOUCH	\\
(Perez et al. 2011)	&	2011	&	JEB	&	mammals	&	29	&	phylogenetic correction	&	APE	\\
(Raia and Meiri 2011)	&	2011	&	Evolution	&	mammals	&	842	&	model of evolution	&	MOTMOT	\\
(Rosas-Guerrero et al. 2011)	&	2011	&	Evolution	&	plants	&	20	&	phylogenetic correction	&	APE	\\
(Setiadi et al. 2011)	&	2011	&	Am Nat	&	amphibians	&	22	&	model of evolution	&	OUCH	\\
(Smith et al. 2011)	&	2011	&	Evolution	&	reptiles	&	15	&	model of evolution	&	GEIGER	\\
(Tulli et al. 2011)	&	2011	&	JEB	&	reptiles	&	29	&	phylogenetic correction	&	?	\\
(Turbill et al. 2011)	&	2011	&	PRSB	&	mammals	&	19	&	phylogenetic correction	&	GEIGER	\\
(Valido et al. 2011)	&	2011	&	JEB	&	plants	&	111	&	phylogenetic correction	&	APE	\\
(Wiens et al. 2011)	&	2011	&	ELE	&	amphibians	&	337	&	model of evolution/	&	GEIGER/	\\
	&		&		&		&		&	phylogenetic correction	&	COMPARE	\\
(Weir and Wheatcroft 2011)	&	2011	&	PRSB	&	birds	&	232	&	model of evolution	&	GEIGER	\\
(Beaulieu et al. 2012)	&	2012	&	Evolution	&	plants	&	590	&	model of evolution	&	OUwie	\\
(Betancur-R et al. 2012)	&	2012	&	ELE	&	fish	&	123	&	model of evolution	&	GEIGER	\\
(Blankers et al. 2012)	&	2012	&	JEB	&	amphibians	&	189	&	phylogenetic correction	&	GEIGER	\\
(Boettiger et al. 2012)	&	2012	&	Evolution	&	reptiles	&	23	&	ancestral state reconstruction	&	OUCH	\\
(Burbrink et al. 2012)	&	2012	&	PRSB	&	multiple	&	41	&	model of evolution	&	Custom code	\\
(Calosi et al. 2012)	&	2012	&	JEB	&	insects	&	25	&	phylogenetic correction	&	MATLAB	\\
(Claramunt et al. 2012a)	&	2012	&	Am Nat	&	birds	&	290	&	model of evolution	&	GEIGER	\\
(Claramunt et al. 2012b)	&	2012	&	PRSB	&	birds	&	282	&	model of evolution	&	GEIGER	\\
(Davis et al. 2012)	&	2012	&	JEB	&	insects	&	53	&	phylogenetic correction	&	SLOUCH	\\
(Diniz-Filho et al. 2012)	&	2012	&	Evolution	&	mammals	&	209	&	other	&	PAM	\\
(Fusco et al. 2012)	&	2012	&	Evolution	&	trilobites	&	60	&	model of evolution	&	?	\\
(Gomez-Mestre et al. 2012)	&	2012	&	Evolution	&	amphibians	&	720	&	phylogenetic correction	&	APE	\\
(Ingram et al. 2012)	&	2012	&	JEB	&	food webs	&	20	&	model of evolution	&	GEIGER	\\
(Kellermann et al. 2012a)	&	2012	&	Evolution	&	insects	&	94	&	phylogenetic correction	&	SLOUCH	\\
(Kellermann et al. 2012b)	&	2012	&	PNAS	&	insects	&	94	&	phylogenetic correction	&	SLOUCH	\\
(Nogueira et al. 2012)	&	2012	&	JEB	&	plants	&	105	&	phylogenetic signal	&	GEIGER	\\
(Ord 2012)	&	2012	&	JEB	&	reptiles	&	32	&	ancestral state reconstruction	&	SLOUCH	\\
(Pearse and Hipp 2012)	&	2012	&	Evolution	&	plants	&	56	&	phylogenetic correction	&	SLOUCH	\\
(Pellissier et al. 2012)	&	2012	&	JEB	&	insects	&	83	&	model of evolution	&	? (APE)	\\
(Price et al. 2012)	&	2012	&	Evolution	&	fish	&	50	&	model of evolution	&	GEIGER	\\
(Sallan and Friedman 2012)	&	2012	&	PRSB	&	fish	&	100	&	model of evolution	&	GEIGER	\\
(Santana et al. 2012)	&	2012	&	Evolution	&	mammals	&	85	&	model of evolution	&	OUCH	\\
(Schmerler et al. 2012)	&	2012	&	PRSB	&	plants	&	88	&	phylogenetic correction	&	nlme	\\
(Smith 2012)	&	2012	&	Evolution	&	birds	&	42	&	phylogenetic correction	&	APE/nlme	\\
(Sookias et al. 2012)	&	2012	&	PRSB	&	multiple	&	43	&	model of evolution	&	GEIGER	\\
(Stireman et al. 2012)	&	2012	&	JEB	&	insects	&	24	&	model of evolution	&	GEIGER	\\
(Voje and Hansen 2012)	&	2012	&	Evolution	&	insects	&	30	&	phylogenetic correction	&	SLOUCH	\\
(Weir et al. 2012)	&	2012	&	Evolution	&	birds	&	232	&	model of evolution	&	GEIGER	\\
(Arbour and López-Fernández 2013)	&	2013	&	PRSB	&	fish	&	27	&	model of evolution	&	OUCH	\\
(Benesh et al. 2013)	&	2013	&	Am Nat	&	helminths	&	143	&	phylogenetic correction	&	APE	\\
(Blackburn et al. 2013)	&	2013	&	Evolution	&	amphibians	&	18	&	model of evolution	&	GEIGER	\\
(Christin et al. 2013)	&	2013	&	PNAS	&	plants	&	545	&	model of evolution	&	GEIGER/OUCH	\\
(Frédérich et al. 2013)	&	2013	&	Am Nat	&	fish	&	208	&	model of evolution	&	OUwie	\\
(Friedman et al. 2013)	&	2013	&	Evolution	&	birds	&	15	&	model of evolution	&	OUCH	\\
(Guerrero et al. 2013)	&	2013	&	PNAS	&	plants/reptiles	&	49	&	ancestral state reconstruction	&	GEIGER/	\\
	&		&		&		&		&		&	COMPARE	\\
(Hertz et al. 2013)	&	2013	&	Evolution	&	reptiles	&	100	&	model of evolution	&	GEIGER	\\
(Hossie et al. 2013)	&	2013	&	JEB	&	amphibians/	&	104	&	phylogenetic correction	&	GEIGER	\\
	&		&		&	reptiles	&		&		&		\\
(Knope and Scales 2013)	&	2013	&	JEB	&	fish	&	26	&	model of evolution	&	OUCH	\\
(Kostikova et al. 2013)	&	2013	&	Am Nat	&	plants	&	68	&	model of evolution	&	OUwie	\\
(Lambert and Wiens 2013)	&	2013	&	Evolution	&	reptiles	&	117	&	ancestral state reconstruction	&	GEIGER	\\
(Lapiedra et al. 2013)	&	2013	&	PRSB	&	birds	&	154	&	model of evolution	&	OUwie	\\
(Litsios et al. 2013)	&	2013	&	Evolution	&	plants	&	382	&	model of evolution	&	OUwie	\\
(López-Fernández et al. 2013)	&	2013	&	Evolution	&	fish	&	135	&	model of evolution	&	GEIGER	\\
(Machac et al. 2013)	&	2013	&	Evolution	&	mammals	&	231	&	model of evolution	&	GEIGER	\\
(Mahler et al. 2013)	&	2013	&	Science	&	reptiles	&	100	&	convergent evolution	&	SURFACE	\\
(Maia et al. 2013)	&	2013	&	PNAS	&	birds	&	47	&	model of evolution	&	OUwie	\\
(Mirceta et al. 2013)	&	2013	&	Science	&	mammals	&	130	&	phylogenetic correction/	&	Mesquite/	\\
	&		&		&		&		&	ancestral state reconstruction	&	MATLAB	\\
(Moen et al. 2013)	&	2013	&	PRSB	&	amphibians	&	44	&	convergent evolution	&	GEIGER	\\
(Pérez i de Lanuza et al. 2013)	&	2013	&	JEB	&	reptiles	&	42	&	ancestral state reconstruction	&	GEIGER	\\
(Pienaar et al. 2013)	&	2013	&	ELE	&	birds	&	382	&	model of evolution	&	SLOUCH	\\
(Quintero and Wiens 2013)	&	2013	&	ELE	&	multiple	&	500	&	ancestral state reconstruction	&	GEIGER/COMPARE	\\
(Ryan and Shaw 2013)	&	2013	&	PRSB	&	mammals	&	34	&	phylogenetic correction	&	MATLAB	\\
(Seddon et al. 2013) 	&	2013	&	PRSB	&	birds	&	153	&	model of evolution	&	? (GEIGER)	\\
(Tanabe and Sota 2013)	&	2013	&	Evolution	&	millipedes	&	84	&	phylogenetic correction	&	APE	\\
(Voje et al. 2013)	&	2013	&	JEB	&	fish	&	87	&	other	&	SLOUCH	\\
(Wiens et al. 2013)	&	2013	&	Evolution	&	reptiles	&	117	&	ancestral state reconstruction	&	GEIGER	\\
\hline
\caption[Literature review]
			{Details of the papers used in our literature review. For each paper we recorded the study taxon, the number of tips in the phylogeny used to fit the Ornstein Uhlenbeck (OU) model, how the authors used the OU model in the paper, and the statistical package (usually R) used to carry out the analyses.For a full reference list see below.}
\label{tab:sk.measurements}

\end{longtable}